---
title: "Final Project Milestone #4"
author: "Julia Englebert"
date: "3/4/2020"
output: html_document
---

## Data Cleaning Progress

My data is sourced from the Inter-university Consortium for Political and Social Research (ICPSR) at icpsr.umich.edu. It was collected for Myron P. Gutmann's Great Plains Population and Environment Data series, and it comes specifically from two subsets of that series: Agricultural Data, 1870-1997 and Social and Demographic Data, 1870-2000. Gutmann conducted this research under the auspices of the University of Michigan.

So far, I have read in all of the data, selecting the variables that are relevant for my project. However, since the data from each year is stored in a unique subdirectory, I will need to bind all of the rows together in order to amass two coherent datasets. Because the variables recorded each year and the methods by which they were measured changed considerably from 1870 to 2000 (the approximate date range of my data), I will first have to modify all of the variables to make them consistent over time. For Milestone 4, I focused on created uniform variables for the agricultural dataset. To name just a few examples, I tried to make adjustments for the shifting categorization of farm yields and the alternating tendencies to report aggregate and average land area in various years.

I plan to continue making the agricultural dataset variables more consistent in the coming weeks, and to repeat a similar process for the demographic dataset. Once this is complete, it should be fairly easy to bind the rows together in order to create a grand total of two dataframes--instead of the 36 that I am currently grappling with. When my two dataframes are complete, I will be able to begin exploring them in greater depth, and I will start trying to understand and visualize the relationship between the two. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(haven)
library(tidyverse)
library(gt)
library(devtools)
library(tibble)
library(foreign)
```

```{r, include=FALSE}
```


```{r, include=FALSE}
# Reading in all of my ag files

# 1870-1997
# Ag data

# Used AFP_XX_V as opposed to STK_XX_V, FRU_XX_V, GDN_XX_V, FST_XX_V
tbl_1 <- read_por("ICPSR_04254_ag/DS0001/04254-0001-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_XX_V, IMP_XX_V, AFP_XX_V))

tbl_2 <- read_por("ICPSR_04254_ag/DS0002/04254-0002-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_XX_V, IMP_XX_V, AFP_XX_V))

tbl_3 <- read_por("ICPSR_04254_ag/DS0003/04254-0003-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_XX_V, IMP_XX_V, AFP_XX_V))

# Some columns are changed past this point
# PRP_XX_V => PRP_OF_V
# IMP_XX_V => IMP_OF_V
# STK_XX_V, FRU_XX_V, GDN_XX_V, FST_XX_V => STK_XX_V, CRO_XX_V (livestock, crop values)

# For tbl_4
# PRP_XX_V => FML_XX_V + BLD_XX_V
# STK_XX_V: Value of Farm Property: Live Stock: Dollars
# AFP_XQ_V: Products not fed to animals

tbl_4 <- read_por("ICPSR_04254_ag/DS0004/04254-0004-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, FML_XX_V, BLD_XX_V, IMP_XX_V, STK_XX_V, AFP_XQ_V))

# Livestock value data is missing for this year tbl_5
tbl_5 <- read_por("ICPSR_04254_ag/DS0005/04254-0005-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_OF_V, IMP_XX_V, CRO_XX_V))

tbl_6 <- read_por("ICPSR_04254_ag/DS0006/04254-0006-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_OF_V, IMP_XX_V, CRO_XX_V, STK_XX_V))

# Total cash value of farm PRP_XX_V missing for tbl_7
tbl_7 <- read_por("ICPSR_04254_ag/DS0007/04254-0007-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, IMP_XX_V, CRO_XX_V, STK_XX_V))

tbl_8 <- read_por("ICPSR_04254_ag/DS0008/04254-0008-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_OF_V, IMP_XX_V, CRO_XX_V, STK_XX_V))

# Limited data for tbl_9
tbl_9 <- read_por("ICPSR_04254_ag/DS0009/04254-0009-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q))

tbl_10 <- read_por("ICPSR_04254_ag/DS0010/04254-0010-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_OF_V, IMP_OF_V, CRO_XX_V, STK_XX_V))

# Limited data for tbl_9
tbl_11 <- read_por("ICPSR_04254_ag/DS0011/04254-0011-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, IMP_XX_V))

# PRP_XX_V => PRP_AX_V * FRM_XX_Q (average value of farm * number of farms)
# Missing IMP_XX_V
# Used AFP_TX_V instead of CRO_XX_V, STK_XX_V
tbl_12 <- read_por("ICPSR_04254_ag/DS0012/04254-0012-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_AX_V, AFP_TX_V))

# PRP_XX_V and IMP_XX_V unknown
# CRO_XX_V, STK_XX_V => AFP_TX_V (all prducts sold)
tbl_13 <- read_por("ICPSR_04254_ag/DS0013/04254-0013-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_SA_Q:FRM_SL_Q, AFP_TX_V))

# PRP_XX_V and IMP_XX_V unknown
# CRO_XX_V, STK_XX_V => AFP_TX_V (all prducts sold)
tbl_14 <- read_por("ICPSR_04254_ag/DS0014/04254-0014-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, AFP_TX_V))

# Same comments as above
# AFP_TX_V = all products sold
tbl_15 <- read_por("ICPSR_04254_ag/DS0015/04254-0015-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, AFP_TX_V))

tbl_16 <- read_por("ICPSR_04254_ag/DS0016/04254-0016-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, AFP_TX_V))

tbl_17 <- read_por("ICPSR_04254_ag/DS0017/04254-0017-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, AFP_TX_V))
tbl_18 <- read_por("ICPSR_04254_ag/DS0018/04254-0018-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, AFP_TX_V))
tbl_19 <- read_por("ICPSR_04254_ag/DS0019/04254-0019-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, AFP_TX_V))
tbl_20 <- read_por("ICPSR_04254_ag/DS0020/04254-0020-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, AFP_TX_V))
tbl_21 <- read_por("ICPSR_04254_ag/DS0021/04254-0021-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, AFP_TX_V))
tbl_22 <- read_por("ICPSR_04254_ag/DS0022/04254-0022-Data.por", col_select = c(NAME, UNFIPS, GP, FML_XX_A, ARE_XX_A, FRM_XX_Q, AFP_TX_V))
```


```{r, include=FALSE}
# Reading in all of demog files
# 1870-2000

tbl_1_d <- read_por("ICPSR_04296_demog/DS0001/04296-0001-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, POP10P, SCH))

# Very little school enrollment data for this year
tbl_2_d <- read_por("ICPSR_04296_demog/DS0002/04296-0002-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29))

# NOTSCH Persons age 3 and over who are enrolled in school
tbl_3_d <- read_por("ICPSR_04296_demog/DS0003/04296-0003-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH))

# No school data
tbl_4_d <- read_por("ICPSR_04296_demog/DS0004/04296-0004-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29))

tbl_5_d <- read_por("ICPSR_04296_demog/DS0005/04296-0005-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH69, SCH1014, SCH1517, SCH1820))

tbl_6_d <- read_por("ICPSR_04296_demog/DS0006/04296-0006-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH713, SCH1415, SCH1617, SCH1820))

tbl_7_d <- read_por("ICPSR_04296_demog/DS0007/04296-0007-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH713, SCH1415, SCH1617, SCH1820))

tbl_8_d <- read_por("ICPSR_04296_demog/DS0008/04296-0008-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH56, SCH713, SCH1415, SCH1617, SCH1820))

tbl_9_d <- read_por("ICPSR_04296_demog/DS0009/04296-0009-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH56, SCH713, SCH1415, SCH1617, SCH1819, SCH2024))

# This year the data is aggragated, sadly
tbl_10_d <- read_por("ICPSR_04296_demog/DS0010/04296-0010-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, ENROLLCO, ENROLLEL, ENROLLHS, SCH534))

tbl_11_d <- read_por("ICPSR_04296_demog/DS0011/04296-0011-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, ENROLLCO, ENROLLEL, ENROLLHS))
tbl_12_d <- read_por("ICPSR_04296_demog/DS0012/04296-0012-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, ENROLLCO, ENROLLEL, ENROLLHS))
tbl_13_d <- read_por("ICPSR_04296_demog/DS0013/04296-0013-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, ENROLLCO, GRHI))
tbl_14_d <- read_por("ICPSR_04296_demog/DS0014/04296-0014-Data.por", col_select = c(TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, ENROLLCO, GRHI))

# Make all the variables the same
# Make sure they also have the same levels
# bind_rows

```


```{r, include=FALSE}

#Notes to myself about variable names

# These are the relevant columns

# Ag data

# FML_XX_A Acres of farms
# ARE_XX_A Total Land Area: Acres
# FRM_XX_Q Number of Farms: Of All Sizes
# PRP_XX_V Present Cash Value: Of Farms: Dollars
# IMP_XX_V Present Cash Value: of Farming Implements and Machinery: Dollars
# FRU_XX_V Orchard Products: Dollars
# GDN_XX_V Produce of Market Gardens: Dollars
# FST_XX_V Forest Products: Dollars
# STK_XX_V Value of livestock

# Changes

# PRP_XX_V => PRP_OF_V
# IMP_XX_V => IMP_OF_V
# STK_XX_V, FRU_XX_V, GDN_XX_V, FST_XX_V => STK_XX_V, CRO_XX_V (livestock, crop values)

# Demog data

# TTOT total persons
# FT0_4 females 0-4
# FT5_9
# FT10_14
# FT15_19
# FT20_24
# FT25_29
# MT0_4 males 0-4
# MT5_9
# MT10_14
# MT15_19
# MT20_24
# MT25_29
# ILL10P Persons age 10 and over who are illiterate
# NWRI10P Persons age 10 and over who cannot write
# POP10P Persons age 10 and over
# SCH Pupils in common schools


# pivot_longer(data = tbl_1_2, cols = )
```